A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Beam Dynamics

### 2018-07-10 paper/paper.tex
updating paper
remove some CS stuff which I think is to much for Phys. Rev. AB

@@

By abiding to the API from Listing 3 it becomes trivial to add new forward
  solvers to the framework (see Appendix A).
If the objectives and constraints are simple arithmetical expressions, the
  \texttt{FunctionEvaluator} simulator can be used.
Using functors and the default expression primitives already powerful
  multi-objective optimization problems can be specified, i.e.\ the benchmark
  problem presented in \cite{hbwh:05}:
%
\begin{align} \label{eqn:bench}
  \text{ min} \quad & \quad \left[ 1 - \exp \left( -1 \left(
    \left(x_1 - \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_2 - \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_3 - \frac{1}{\sqrt{3}} \right)^2 \right)\right), \right. \\
                    & \quad \left. 1 - \exp \left( -1 \left(
    \left(x_1 + \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_2 + \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_3 + \frac{1}{\sqrt{3}} \right)^2 \right)\right) \right]^T \nonumber \\
    \text{s.t.} \quad & \quad -1 \le x_i \le 1, \quad i=1,2,3 \nonumber
  \text{.}
\end{align}
%
Using the default expression primitives this can be stated in an input file
  as:

@@

\input{theor-experiment}

@@

The authors thank the \textsc{SwissFEL} and AWA teams for contributing to the
  formulation of optimization problems. 
  We gratefully acknowledge the computing resources provided on Bebop,
  a HPC cluster operated by the LCRC at ANL.
  This work was partly supported by the 
  U.S. Department of Energy, Office of Science, under 
  contract number DE-AC02-06CH11357 and grant number DE-SC0015479. 
\appendix

\section{Forward Solver Implementation}

A simple implementation, e.g.\ using a Python forward solver, is given below
  in Listing 6.
We assume the Python script executes a simulation and dumps results in the
  \textsc{SDDS} file format~\cite{borl:00}.
Utility classed are used to parse the result data and fill the appropriate
  structures returned to the optimizer algorithm.

\begin{code}[Simple Simulation Wrapper]
#include <map>
#include <string>
#include <vector>

#include "Util/Types.h"
#include "Util/CmdArguments.h"
#include "Simulation/Simulation.h"

#include "boost/smart_ptr.hpp"

class SimpleSimulation : public Simulation {

public:

    SimpleSimulation(Expressions::Named_t objectives,
                     Expressions::Named_t constraints,
                     Param_t params, std::string name, MPI_Comm comm,
                     CmdArguments_t args);

    ~SimpleSimulation();

    /// Calls simulation through Python wrapper and returns when simulation
    /// has either failed or finished.
    void run() {
        // user implements wrappers
        prepare_input_file();
        run_python_simulation();
    }

    /// Parse SDDS stat file and build up requested variable dictionary.
    void collectResults(){

        Expressions::Named_t::iterator it;
        for(it = objectives_.begin(); it != objectives_.end(); it++) {

            Expressions::Expr_t *objective = it->second;

            // find out which variables we need in order to evaluate the
            // objective
            variableDictionary_t variable_dictionary;
            std::set<std::string> req_vars = objective->getReqVars();

            if(req_vars.size() != 0) {

                boost::scoped_ptr<SDDSReader> sddsr(new SDDSReader(fn));

                try {
                    sddsr->parseFile();
                } catch(OptPilotException &e) {
                    std::cout << "Exception while parsing SDDS file: "
                        << e.what() << std::endl;
                    break;
                }

                // get all the required variable values from the stat file
                foreach(std::string req_var, req_vars) {
                    if(variable_dictionary.count(req_var) == 0) {

                        try {
                            double value = 0.0;
                            sddsr->getValue(1 /*atTime*/, req_var, value);
                            variable_dictionary.insert(
                                std::pair<std::string, double>(req_var, value));
                        } catch(OptPilotException &e) {
                            std::cout << "Exception while getting value "
                                      << "from SDDS file: " << e.what()
                                      << std::endl;
                        }
                    }
                }
            }

            // and evaluate the expression using the built dictionary of
            // variable values
            Expressions::Result_t result =
                objective->evaluate(variable_dictionary);

            std::vector<double> values;
            values.push_back(boost::get<0>(result));
            bool is_valid = boost::get<1>(result);

            reqVarInfo_t tmps = {EVALUATE, values, is_valid};
            requestedVars_.insert(
                std::pair<std::string, reqVarInfo_t>(it->first, tmps));

        }
    }

    /// returns container containing all requested variables with results
    reqVarContainer_t getResults() { return requestedVars_; }

private:

    /// holds solutions returned to the optimizer
    reqVarContainer_t requestedVars_;

    Expressions::Named_t objectives_;
    Expressions::Named_t constraints_;

    MPI_Comm comm_;
};

#endif
\end{code}

>>>>

\usepackage{graphicx,xcolor,enumitem}
\newcommand{\lsnote}[1]{\textsf{{\color{violet}{ LS note:}   #1 }}}

@@

% AA: this is too much for Phys Rev. AB  
% By abiding to the API from Listing 3 it becomes trivial to add new forward
% solvers to the framework (see Appendix A).

If the objectives and constraints are simple arithmetical expressions, the
  \texttt{FunctionEvaluator} simulator can be used.
Using functors and the default expression primitives already powerful
  multi-objective optimization problems can be specified, i.e.\ the benchmark
  problem presented in \cite{hbwh:05}:
%
\begin{align} \label{eqn:bench}
  \text{ min} \quad & \quad \left[ 1 - \exp \left( -1 \left(
    \left(x_1 - \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_2 - \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_3 - \frac{1}{\sqrt{3}} \right)^2 \right)\right), \right. \\
                    & \quad \left. 1 - \exp \left( -1 \left(
    \left(x_1 + \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_2 + \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_3 + \frac{1}{\sqrt{3}} \right)^2 \right)\right) \right]^T \nonumber \\
    \text{s.t.} \quad & \quad -1 \le x_i \le 1, \quad i=1,2,3 \nonumber
  \text{.}
\end{align}
%
Using the default expression primitives this can be stated in an input file
  as:

@@

\input{theor-experiment} %periment}

@@

The authors thank the \textsc{SwissFEL} and AWA teams for contributing to the
  formulation of optimization problems. 
  We gratefully acknowledge the computing resources provided on Bebop,
  a HPC cluster operated by the LCRC at ANL.
  This work was partly supported by the 
  U.S. Department of Energy, Office of Science, under 
  contract number DE-AC02-06CH11357 and grant number DE-SC0015479.

% AA: this is too much for Phys Rev. AB

% \appendix

%\section{Forward Solver Implementation}
%
%A simple implementation, e.g.\ using a Python forward solver, is given below
%  in Listing 6.
%We assume the Python script executes a simulation and dumps results in the
%  \textsc{SDDS} file format~\cite{borl:00}.
%Utility classed are used to parse the result data and fill the appropriate
%  structures returned to the optimizer algorithm.
%
%\begin{code}[Simple Simulation Wrapper]
%#include <map>
%#include <string>
%#include <vector>
%
%#include "Util/Types.h"
%#include "Util/CmdArguments.h"
%#include "Simulation/Simulation.h"
%
%#include "boost/smart_ptr.hpp"
%
%class SimpleSimulation : public Simulation {
%
%public:
%
%    SimpleSimulation(Expressions::Named_t objectives,
%                     Expressions::Named_t constraints,
%                     Param_t params, std::string name, MPI_Comm comm,
%                     CmdArguments_t args);
%
%    ~SimpleSimulation();
%
%    /// Calls simulation through Python wrapper and returns when simulation
%    /// has either failed or finished.
%    void run() {
%        // user implements wrappers
%        prepare_input_file();
%        run_python_simulation();
%    }
%
%    /// Parse SDDS stat file and build up requested variable dictionary.
%    void collectResults(){
%
%        Expressions::Named_t::iterator it;
%        for(it = objectives_.begin(); it != objectives_.end(); it++) {
%
%            Expressions::Expr_t *objective = it->second;
%
%            // find out which variables we need in order to evaluate the
%            // objective
%            variableDictionary_t variable_dictionary;
%            std::set<std::string> req_vars = objective->getReqVars();
%
%            if(req_vars.size() != 0) {
%
%                boost::scoped_ptr<SDDSReader> sddsr(new SDDSReader(fn));
%
%                try {
%                    sddsr->parseFile();
%                } catch(OptPilotException &e) {
%                    std::cout << "Exception while parsing SDDS file: "
%                        << e.what() << std::endl;
%                    break;
%                }
%
%                // get all the required variable values from the stat file
%                foreach(std::string req_var, req_vars) {
%                    if(variable_dictionary.count(req_var) == 0) {
%
%                        try {
%                            double value = 0.0;
%                            sddsr->getValue(1 /*atTime*/, req_var, value);
%                            variable_dictionary.insert(
%                                std::pair<std::string, double>(req_var, value));
%                        } catch(OptPilotException &e) {
%                            std::cout << "Exception while getting value "
%                                      << "from SDDS file: " << e.what()
%                                      << std::endl;
%                        }
%                    }
%                }
%            }
%
%            // and evaluate the expression using the built dictionary of
%            // variable values
%            Expressions::Result_t result =
%                objective->evaluate(variable_dictionary);
%
%            std::vector<double> values;
%            values.push_back(boost::get<0>(result));
%            bool is_valid = boost::get<1>(result);
%
%            reqVarInfo_t tmps = {EVALUATE, values, is_valid};
%            requestedVars_.insert(
%                std::pair<std::string, reqVarInfo_t>(it->first, tmps));
%
%        }
%    }
%
%    /// returns container containing all requested variables with results
%    reqVarContainer_t getResults() { return requestedVars_; }
%
%private:
%
%    /// holds solutions returned to the optimizer
%    reqVarContainer_t requestedVars_;
%
%    Expressions::Named_t objectives_;
%    Expressions::Named_t constraints_;
%
%    MPI_Comm comm_;
%};
%
%#endif
%\end{code}
